---
title: "Structural Equation Models 2019"
output:
  html_document: default
---

## Week 4


### Exercise 4.1

Let us import the data file in R, prepare the data and build the model according to the input file. Then let us draw the graph of the initial, full structural equation model.

```{r echo=FALSE,warning=FALSE}
wk4 <- read.table(file = "allsecondary.DAT")

colnames(wk4) <- c("ROLEA1", "ROLEA2", "ROLEC1", "ROLEC2", "WORK1", "WORK2",
                   "CCLIM1", "CCLIM2", "CCLIM3", "CCLIM4", "DEC1", "DEC2",
                   "SSUP1", "SSUP2", "PSUP1", "PSUP2", "SELF1", "SELF2",
                   "SELF3", "ELC1", "ELC2", "ELC3", "ELC4", "ELC5",
                   "EE1", "EE2", "EE3", "DP1", "DP2", "PA1", "PA2", "PA3")

library(lavaan)

model <- '
# latent variables
F1  =~ ROLEA1 + ROLEA2 + DEC2
F2  =~ ROLEC1 + ROLEC2
F3  =~ WORK1 + WORK2
F4  =~ CCLIM1 + CCLIM2 + CCLIM3 + CCLIM4
F5  =~ DEC1 + DEC2
F6  =~ SSUP1 + SSUP2 + DEC2
F7  =~ PSUP1 + PSUP2
F8  =~ SELF1 + SELF2 + SELF3
F9  =~ ELC1 + ELC2 + ELC3 + ELC4 + ELC5
F10  =~ EE1 + EE2 + EE3
F11  =~ DP1 + DP2
F12  =~ PA1 + PA2 + PA3

# regressions
F8 ~ F5 + F6 + F7
F9 ~ F5
F10 ~ F2 + F3 + F4 
F11 ~ F2 + F10
F12 ~ F1 + F8 + F9 + F10 + F11
'

fit <- sem(model, data = wk4, estimator="MLM")

library(semPlot)

semPaths(fit, what = "path", whatLabels = "hide", style = "lisrel",
         layout = "tree2", residuals = TRUE, nCharNodes = 0, rotation = 2, 
         sizeMan = 6, sizeMan2= 4, sizeLat = 7, esize = 1, label.cex = 1.2,
         optimizeLatRes = TRUE, exoCov = FALSE)
```



Then let us estimate the initial model using the robust MLM estimator and present a brief summary of the model fit:



```{r echo=FALSE,warning=FALSE}
# summary(fit, fit.measures = T)
library(dplyr)
msr <- c("chisq.scaled", "df.scaled", "pvalue.scaled", 
         "chisq.scaling.factor", "cfi.robust", "tli.robust",
         "rmsea.robust")

M1 <- c(fitMeasures(fit)[msr]) %>% round(3)
print(M1) 
```



In the output Chi square statistic 1541.844 (df = 427, p = 0.000) suggests against a good fit. However the fit indices robust CFI (0.945), robust TLI (0.936) and robust RMSEA (0.045) indicate good fit. 



### Exercise 4.2


Then let us print the modification indices (MI) for the model:

```{r echo=FALSE,warning=FALSE}
modindices(fit, minimum.value = 60, sort = TRUE, op="~")
```

#### Model 2
F11 ~ F4 has the highest MI value. Following the lecture material we specify another model by adding F11 ~ F4:


```{r echo=FALSE,warning=FALSE}
model2 <- '
# latent variables
F1  =~ ROLEA1 + ROLEA2 + DEC2
F2  =~ ROLEC1 + ROLEC2
F3  =~ WORK1 + WORK2
F4  =~ CCLIM1 + CCLIM2 + CCLIM3 + CCLIM4
F5  =~ DEC1 + DEC2
F6  =~ SSUP1 + SSUP2 + DEC2
F7  =~ PSUP1 + PSUP2
F8  =~ SELF1 + SELF2 + SELF3
F9  =~ ELC1 + ELC2 + ELC3 + ELC4 + ELC5
F10  =~ EE1 + EE2 + EE3
F11  =~ DP1 + DP2
F12  =~ PA1 + PA2 + PA3

# regressions
F8 ~ F5 + F6 + F7
F9 ~ F5
F10 ~ F2 + F3 + F4 
F11 ~ F2 + F10 + F4
F12 ~ F1 + F8 + F9 + F10 + F11
'

fit2 <- sem(model2, data = wk4, estimator = "MLM")
M2 <- c(fitMeasures(fit2)[msr]) %>% round(3) 
print(M2) 
```

Let us list the indices from two first models so they can be compared:

```{r echo=FALSE,warning=FALSE}
tbl1 <- matrix(c(M1, M2), nrow=2, byrow = T)
nms <- c("ChiSquare", "DF", "p", "ChiSq.sc.factor", "CFI", "TLI", "RMSEA")
colnames(tbl1) <- nms
rownames(tbl1) <- c("M1", "M2")
tbl1
```

The model 2 seems to have better fit than the original model: Chi square statistic is still significant 1440.864 (df = 426, p = 0.000) and suggests against a good fit. However the fit indices robust CFI (0.95), robust TLI (0.942) and robust RMSEA (0.043) all indicate slightly better fit than original model.


Then let us calculate the Chi-Square Difference Test:
```{r echo=FALSE,warning=FALSE}
cd <- ((427*1.127)-(426*1.125))/(427-426)
TRd <- (1541.844*1.127 - 1440.864*1.125)/cd
TRd 
```


```{r echo=FALSE,warning=FALSE}
anova(fit2, fit)
```

Both tests show almost exactly the same results and indicate significant difference between models.


#### Model 3

Next, we specify a third model by adding F12 ~ F5:


```{r echo=FALSE,warning=FALSE}
model3 <- '
# latent variables
F1  =~ ROLEA1 + ROLEA2 + DEC2
F2  =~ ROLEC1 + ROLEC2
F3  =~ WORK1 + WORK2
F4  =~ CCLIM1 + CCLIM2 + CCLIM3 + CCLIM4
F5  =~ DEC1 + DEC2
F6  =~ SSUP1 + SSUP2 + DEC2
F7  =~ PSUP1 + PSUP2
F8  =~ SELF1 + SELF2 + SELF3
F9  =~ ELC1 + ELC2 + ELC3 + ELC4 + ELC5
F10  =~ EE1 + EE2 + EE3
F11  =~ DP1 + DP2
F12  =~ PA1 + PA2 + PA3

# regressions
F8 ~ F5 + F6 + F7
F9 ~ F5
F10 ~ F2 + F3 + F4 
F11 ~ F2 + F10 + F4
F12 ~ F1 + F8 + F9 + F10 + F11 + F5
'

fit3 <- sem(model3, data = wk4, estimator = "MLM")
M3 <- c(fitMeasures(fit3)[msr]) %>% round(3) 
print(M3) 
```

```{r echo=FALSE,warning=FALSE}
tbl1 <- rbind(tbl1, M3)
rownames(tbl1) <- c("M1", "M2", "M3")
tbl1
```

The third model seems to improve the fit from the 2nd model slightly in the same manner as the second model improved the original model. 

```{r echo=FALSE,warning=FALSE}
cd2 = (426*1.125 - 425*1.125)/(426 - 425)
TRd2 = (1440.864*1.125 - 1396.527*1.125)/cd2
TRd2
```

The Chi-Square Difference Test indicates that the difference between models is significant.

#### Model 4

Next, we specify a fourth model by adding F9 ~ F2:


```{r echo=FALSE,warning=FALSE}
model4 <- '
# latent variables
F1  =~ ROLEA1 + ROLEA2 + DEC2
F2  =~ ROLEC1 + ROLEC2
F3  =~ WORK1 + WORK2
F4  =~ CCLIM1 + CCLIM2 + CCLIM3 + CCLIM4
F5  =~ DEC1 + DEC2
F6  =~ SSUP1 + SSUP2 + DEC2
F7  =~ PSUP1 + PSUP2
F8  =~ SELF1 + SELF2 + SELF3
F9  =~ ELC1 + ELC2 + ELC3 + ELC4 + ELC5
F10  =~ EE1 + EE2 + EE3
F11  =~ DP1 + DP2
F12  =~ PA1 + PA2 + PA3

# regressions
F8 ~ F5 + F6 + F7
F9 ~ F5 + F2
F10 ~ F2 + F3 + F4 
F11 ~ F2 + F10 + F4
F12 ~ F1 + F8 + F9 + F10 + F11 + F5
'

fit4 <- sem(model4, data = wk4, estimator = "MLM")
M4 <- c(fitMeasures(fit4)[msr]) %>% round(3) 
print(M4) 
```

```{r echo=FALSE,warning=FALSE}
tbl1 <- rbind(tbl1, M4)
rownames(tbl1) <- c("M1", "M2", "M3", "M4")
tbl1
```

The fourth model seems to improve the fit from the third model slightly in the same manner as the third model improved the second model. 


#### Model 5

Next, we specify a fifth model by adding F9 ~ F8:


```{r echo=FALSE,warning=FALSE}
model5 <- '
# latent variables
F1  =~ ROLEA1 + ROLEA2 + DEC2
F2  =~ ROLEC1 + ROLEC2
F3  =~ WORK1 + WORK2
F4  =~ CCLIM1 + CCLIM2 + CCLIM3 + CCLIM4
F5  =~ DEC1 + DEC2
F6  =~ SSUP1 + SSUP2 + DEC2
F7  =~ PSUP1 + PSUP2
F8  =~ SELF1 + SELF2 + SELF3
F9  =~ ELC1 + ELC2 + ELC3 + ELC4 + ELC5
F10  =~ EE1 + EE2 + EE3
F11  =~ DP1 + DP2
F12  =~ PA1 + PA2 + PA3

# regressions
F8 ~ F5 + F6 + F7
F9 ~ F5 + F2 + F8
F10 ~ F2 + F3 + F4 
F11 ~ F2 + F10 + F4
F12 ~ F1 + F8 + F9 + F10 + F11 + F5
'

fit5 <- sem(model5, data = wk4, estimator = "MLM")
M5 <- c(fitMeasures(fit5)[msr]) %>% round(3)
```

```{r echo=FALSE,warning=FALSE}
tbl1 <- rbind(tbl1, M5)
rownames(tbl1) <- c("M1", "M2", "M3", "M4", "M5")
tbl1
```

The fifth model seems to improve the fit from the fourth model slightly in the same manner as the fourth model improved the third model. 


#### Model 6

Next, we specify a sixth model by adding F10 ~ F8:


```{r echo=FALSE,warning=FALSE}
model6 <- '
# latent variables
F1  =~ ROLEA1 + ROLEA2 + DEC2
F2  =~ ROLEC1 + ROLEC2
F3  =~ WORK1 + WORK2
F4  =~ CCLIM1 + CCLIM2 + CCLIM3 + CCLIM4
F5  =~ DEC1 + DEC2
F6  =~ SSUP1 + SSUP2 + DEC2
F7  =~ PSUP1 + PSUP2
F8  =~ SELF1 + SELF2 + SELF3
F9  =~ ELC1 + ELC2 + ELC3 + ELC4 + ELC5
F10  =~ EE1 + EE2 + EE3
F11  =~ DP1 + DP2
F12  =~ PA1 + PA2 + PA3

# regressions
F8 ~ F5 + F6 + F7
F9 ~ F5 + F2 + F8
F10 ~ F2 + F3 + F4 + F8
F11 ~ F2 + F10 + F4
F12 ~ F1 + F8 + F9 + F10 + F11 + F5
'

fit6 <- sem(model6, data = wk4, estimator = "MLM")
M6 <- c(fitMeasures(fit6)[msr]) %>% round(3)
```

```{r echo=FALSE,warning=FALSE}
tbl1 <- rbind(tbl1, M6)
rownames(tbl1) <- c("M1", "M2", "M3", "M4", "M5", "M6")
tbl1
```

The sixth model seems to improve the fit from the fifth model slightly in the same manner as the fifth model improved the fourth model. 

#### Model 7

Next, we specify a seventh model by adding F12 ~ F2:


```{r echo=FALSE,warning=FALSE}
model7 <- '
# latent variables
F1  =~ ROLEA1 + ROLEA2 + DEC2
F2  =~ ROLEC1 + ROLEC2
F3  =~ WORK1 + WORK2
F4  =~ CCLIM1 + CCLIM2 + CCLIM3 + CCLIM4
F5  =~ DEC1 + DEC2
F6  =~ SSUP1 + SSUP2 + DEC2
F7  =~ PSUP1 + PSUP2
F8  =~ SELF1 + SELF2 + SELF3
F9  =~ ELC1 + ELC2 + ELC3 + ELC4 + ELC5
F10  =~ EE1 + EE2 + EE3
F11  =~ DP1 + DP2
F12  =~ PA1 + PA2 + PA3

# regressions
F8 ~ F5 + F6 + F7
F9 ~ F5 + F2 + F8
F10 ~ F2 + F3 + F4 + F8
F11 ~ F2 + F10 + F4
F12 ~ F1 + F8 + F9 + F10 + F11 + F5 + F2
'

fit7 <- sem(model7, data = wk4, estimator = "MLM")
M7 <- c(fitMeasures(fit7)[msr]) %>% round(3)
```

```{r echo=FALSE,warning=FALSE}
tbl1 <- rbind(tbl1, M7)
rownames(tbl1) <- c("M1", "M2", "M3", "M4", "M5", "M6", "M7")
tbl1
```

The seventh model seems to improve the fit from the sixth model slightly in the same manner as the sixth model improved the fifth model. 


#### Model 8

The final model 8 is specified by removing F8 ~ F7 and F12 ~ F1:


```{r echo=FALSE,warning=FALSE}
model8 <- '
# latent variables
F1  =~ ROLEA1 + ROLEA2 + DEC2
F2  =~ ROLEC1 + ROLEC2
F3  =~ WORK1 + WORK2
F4  =~ CCLIM1 + CCLIM2 + CCLIM3 + CCLIM4
F5  =~ DEC1 + DEC2
F6  =~ SSUP1 + SSUP2 + DEC2
F7  =~ PSUP1 + PSUP2
F8  =~ SELF1 + SELF2 + SELF3
F9  =~ ELC1 + ELC2 + ELC3 + ELC4 + ELC5
F10  =~ EE1 + EE2 + EE3
F11  =~ DP1 + DP2
F12  =~ PA1 + PA2 + PA3

# regressions
F8 ~ F5 + F6
F9 ~ F5 + F2 + F8
F10 ~ F2 + F3 + F4 + F8
F11 ~ F2 + F10 + F4
F12 ~ F8 + F9 + F10 + F11 + F5 + F2
'

fit8 <- sem(model8, data = wk4, estimator = "MLM")
M8 <- c(fitMeasures(fit8)[msr]) %>% round(3)
```

```{r echo=FALSE,warning=FALSE}
tbl1 <- rbind(tbl1, M8)
rownames(tbl1) <- c("M1", "M2", "M3", "M4", "M5", "M6", "M7", "M8")
tbl1
```

The final model seems to not improve the fit from the seventh model. All in all all the steps have improved the fit nicely: the Chi Square statistic decreased from 1541.844 to 1275.427, CFI increased from 0.945 to 0.958, TLI increased from 0.936 to 0.951 and RMSEA decreased from 0.045 to 0.040. 

### Exercise 4.3

Let us draw the graphs of the original and the final models:


```{r echo=FALSE,warning=FALSE}
semPaths(fit, what = "std", style = "lisrel", optimizeLatRes = TRUE,
         layout = "tree2", residuals = FALSE, nCharNodes = 0, rotation = 2, 
         sizeMan = 6, sizeMan2= 4, sizeLat = 7, esize = 1, label.cex = 1.2,
         exoCov = FALSE)

semPaths(fit8, what = "std", style = "lisrel", optimizeLatRes = TRUE,
         layout = "tree2", residuals = FALSE, nCharNodes = 0, rotation = 2, 
         sizeMan = 6, sizeMan2= 4, sizeLat = 7, esize = 1, label.cex = 1.2,
         exoCov = FALSE)
```
